---
title: "NYC Precinct Murder + Shootings + Crimes DiD + Augmented Synthetic Control Analysis"
output: github_document
---

## NOTE: for precinct by precinct analysis, see `by_precinct` folder

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

library(tidyverse)
library(augsynth)
library(did)
library(readxl)

crimes_final <- readRDS("data/crimes_final.rds") |>
  filter(!is.na(case)) |>
  filter(PCT %in% c( "41", "42", "44", "48", "52", "25", "73",
                    "60", "67", "69", "70", "71", "101", "105", "113",
                    "43", "47", "49")) |> # cases
  rowwise() |>
   mutate(crime_total = sum(murder, robbery, rape, assault, na.rm = TRUE))
```

## Difference in Difference Analysis

- Use shooting data (precincts 43, 47, 49) from 2004-2024 to conduct a difference in difference analysis

```{r}
dat49 <- read_excel("data/SUV vs NYPD PCNT DATA_2004_2024.xlsx", sheet = 1) |>
  janitor::clean_names() |>
  mutate(across(everything(), ~ as.numeric(gsub("[^0-9.-]", "", .)))) |>
  mutate(PCT = 49)

dat47 <- read_excel("data/SUV vs NYPD PCNT DATA_2004_2024.xlsx", sheet = 2) |>
  janitor::clean_names() |>
  mutate(across(everything(), ~ as.numeric(gsub("[^0-9.-]", "", .)))) |>
  mutate(PCT = 47)

dat43 <- read_excel("data/SUV vs NYPD PCNT DATA_2004_2024.xlsx", sheet = 3) |>
  janitor::clean_names() |>
  mutate(across(everything(), ~ as.numeric(gsub("[^0-9.-]", "", .)))) |>
  mutate(PCT = 43)

merged_dat <- dat49 |>
  merge(dat47, all = TRUE) |>
  merge(dat43, all = TRUE) |>
  pivot_longer(cols = c(suv_target_area, pcnt_other_areas), 
               names_to = "group", 
               values_to = "y") |>
  mutate(log_y = log(y)) |>
  mutate(treated = ifelse(group == "suv_target_area", 1, 0)) |>
  mutate(treated_year = ifelse(treated == 1, 2015, 0)) |>
  arrange(treated, PCT, year) |>
  mutate(GROUP = as.numeric(paste0(PCT, treated)))

ggplot(merged_dat, aes(x = year, y = log_y, color = factor(group), group = factor(group))) +
  geom_line() +
  geom_point() +
  facet_wrap(~ PCT) +
  geom_vline(xintercept = 2014, color = "black", linetype = "dotted") +
  theme_minimal() +
  labs(title = "log shootings by precinct",
       x = "Year",
       y = "log(shootings)",
       color = "Group")
```

### Removing 2014
```{r}
# removing 2014
set.seed(10)
res <- att_gt(yname = "log_y",
              tname = "year",
              idname = "GROUP",
              #clustervars = c("PCT", "GROUP"),
              gname = "treated_year",
              data = merged_dat  |> filter(year != 2014), # remove 2014 due to rollout period
              control_group = "notyettreated" # includes both those that are never treated and those yet to be treated
)

# aggregate
res_dynamic <- aggte(res, type = "dynamic")
summary(res_dynamic)
ggdid(res_dynamic)
```

### Anticipating 2014
```{r}
# anticipating 2014
set.seed(10)
res_anticipation <- att_gt(yname = "log_y",
                           tname = "year",
                           idname = "GROUP",
                           #clustervars = c("PCT", "GROUP"),
                           gname = "treated_year",
                           data = merged_dat,
                           anticipation = 1,
                           control_group = "notyettreated" # includes both those that are never treated and those yet to be treated
)

# aggregate
res_dynamic_anticipation <- aggte(res_anticipation, type = "dynamic")
summary(res_dynamic_anticipation)
ggdid(res_dynamic_anticipation)

# questions: clustering?
```

## Synthetic Control Analysis (new data)

- Only look at SUV areas of precints 43, 47, 39 for years 2006-2024
- Use other precincts as controls (see below for slection method)

### SHOOTINGS

#### Using ridge to allow for extraploation (lambda selected through a visual analysis)
```{r}
temp <- merged_dat |>
  filter(treated == 1) |>
  filter(year >= 2006) |>
  select(PCT, year, y, treated) |>
  mutate(treated = ifelse(year <= 2014, 0, 1)) |>
  rename("shootings" = "y",
         "case" = "treated") |>
  mutate(PCT = as.character(PCT)) |>
  as.data.frame()


shootings_dat <- crimes_final |>
  filter(!(PCT %in% c(43, 47, 49))) |>
  filter(year >= 2006) |>
  select(PCT, year, shootings, case) |>
  merge(temp, all = TRUE) |>
  mutate(log_shootings = log(shootings)) |>
  filter(year != 2014)

set.seed(5)

# SHOOTINGS
syn_shootings_new <- augsynth(log_shootings ~ case, 
                       unit = PCT, 
                       time = year,
                       data = shootings_dat,
                       progfunc = "Ridge", 
                       scm = T, 
                       lambda = 0.03, 
                       fixedeff = T)

syn_shootings_summ_new <- summary(syn_shootings_new, inf_type = "jackknife+")
syn_shootings_summ_new
plot(syn_shootings_summ_new)

# plots cross-validation
#plot(syn_shootings, cv = T)
```

#### Not allowing for extrapolation (basic model)
```{r}
set.seed(5)

# SHOOTINGS (non extrapolate)
syn_shootings_basic_new <- augsynth(log_shootings ~ case, 
                       unit = PCT, 
                       time = year,
                       data = shootings_dat,
                       progfunc = "None", 
                       scm = T, 
                       fixedeff = T)

syn_shootings_summ_basic_new <- summary(syn_shootings_basic_new, inf_type = "jackknife+")
syn_shootings_summ_basic_new
plot(syn_shootings_summ_basic_new)
```

## Synthetic Control Analysis (old)

### Selecting Precincts:
- Cases: precincts 43, 47, 49
- Controls: chosen based on similar trends/patterns in shootings and murders in the pre-treatment period
  - Bronx: precincts 41, 42, 44, 52
  - Manhattan: precincts 25
  - Brooklyn: precincts 69, 70, 71, 73
  - Queens: precincts 101, 105, 113
- Excluding data in 2014 due to rollout of treatment

### Exploratory Plots
```{r}
hist(crimes_final$murder)
hist(crimes_final$shootings)
hist(crimes_final$crime_total)

crimes_final <- crimes_final |>
  mutate(log_murder = log(murder + 1), #includes a 0 in PCT 60 in 2018
         log_shootings = log(shootings), 
         log_crime_total = log(crime_total)
  )


hist(crimes_final$log_murder)
hist(crimes_final$log_shootings)
hist(crimes_final$log_crime_total)
```

### MURDER

#### Using ridge to allow for extrapolation (lambda selected through a visual analysis)
```{r}
set.seed(5)
# MURDER
syn_murder <- augsynth(log_murder ~ case, 
                       unit = PCT, 
                       time = year,
                       data = crimes_final,
                       progfunc = "Ridge", 
                       scm = T, 
                       lambda = 0.2, 
                       fixedeff = T)

syn_muder_summ <- summary(syn_murder, inf_type = "jackknife+")
syn_muder_summ
plot(syn_muder_summ)
```

### Not allowing for extrapolation (basic model)
```{r}
set.seed(5)
# MURDER (non extrapolate)
syn_murder <- augsynth(log_murder ~ case, 
                       unit = PCT, 
                       time = year,
                       data = crimes_final,
                       progfunc = "None", 
                       scm = T, 
                       fixedeff = T)

syn_muder_summ <- summary(syn_murder, inf_type = "jackknife+")
syn_muder_summ
plot(syn_muder_summ)
```

### SHOOTINGS

#### Using ridge to allow for extraploation (lambda selected through a visual analysis)
```{r}
set.seed(5)

# SHOOTINGS
syn_shootings <- augsynth(log_shootings ~ case, 
                       unit = PCT, 
                       time = year,
                       data = crimes_final |> filter(year >= 2006),
                       progfunc = "Ridge", 
                       scm = T, 
                       lambda = 0.03, 
                       fixedeff = T)

syn_shootings_summ <- summary(syn_shootings, inf_type = "jackknife+")
syn_shootings_summ
plot(syn_shootings_summ)

# plots cross-validation
#plot(syn_shootings, cv = T)
```

#### Not allowing for extrapolation (basic model)
```{r}
set.seed(5)

# SHOOTINGS (non extrapolate)
syn_shootings_basic <- augsynth(log_shootings ~ case, 
                          unit = PCT, 
                          time = year,
                          data = crimes_final |> filter(year >= 2006),
                          progfunc = "None", 
                          scm = T, 
                          fixedeff = T)

syn_shootings_summ_basic <- summary(syn_shootings_basic, inf_type = "jackknife+")
syn_shootings_summ_basic
plot(syn_shootings_summ_basic)
```

### CRIMES (Murder + Rape + Assault + Robbery) -- NOTE: did not include shootings because these may be double counted in murder

#### Using ridge to allow for extrapolation (lambda selected through a visual analysis)
```{r}
set.seed(5)

syn_crime_total <- augsynth(log_crime_total ~ case, 
                            unit = PCT, 
                            time = year,
                            data = crimes_final,
                            progfunc = "Ridge", 
                            scm = T, 
                            lambda = 0.005, 
                            fixedeff = T)

syn_crime_total_summ <- summary(syn_crime_total, inf_type = "jackknife+")
syn_crime_total_summ
plot(syn_crime_total_summ)

# plots cross-validation
#plot(syn_crime_total, cv = T)
```

#### Not allowing for extrapolation (basic model)

```{r}
set.seed(5)

# crime_total (non extrapolate)
syn_crime_total_basic <- augsynth(log_crime_total ~ case, 
                                  unit = PCT, 
                                  time = year,
                                  data = crimes_final,
                                  progfunc = "None", 
                                  scm = T, 
                                  fixedeff = T)

syn_crime_total_summ_basic <- summary(syn_crime_total_basic, inf_type = "jackknife+")
syn_crime_total_summ_basic
plot(syn_crime_total_summ_basic)
```

### Conclusion:
- Treatments may have temporarily reduced shootings in the immediate period following roll-out (until 2016) but there is little evidence to suggest that this has had a lasting impact in reducing shootings when comparing to similar precincts
- There is little evidence that these treatments lowered crime compared to similar precincts, and in fact, there may be evidence to suggest that crime has increased in these 3 precincts compared to other areas

